%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   Filename    : chapter_2.tex 
%
%   Description : This file will contain your review of related works.
%                 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Related Works}
\label{sec:relatedworks}

\section{Test}

\section{AnitoPlume}
AnitoPlume is the first interactive, real-time 3D graphics-based simulation application explicitly tailored for visualizing the changing landscape of the Taal Volcano island in the Philippines and its corresponding eruption plumes.
Developed primarily to address the challenges of volcanic hazard and crisis communication, the simulator serves as an immersive, lightweight alternative to traditional 2D heatmaps and highly complex, non-interactive geoscientific numerical solvers (such as FALL3D and Tephra2).
While geoscientific solvers excel at quantitative predictive accuracy, they require heavy high-performance computing resources and significant processing time, making them unviable for applications requiring immediate visual feedback, interactive flexibility, and low-latency user immersion.
AnitoPlume fills this gap by merging computer graphics techniques with site-specific real-world terrain details to facilitate risk awareness and "what-if" situational modeling \cite{del2026interactive}.

\subsection{Graphics Architecture and System Performance}
To bypass the overhead, heavy abstractions, and unneeded general-purpose features of commercial game engines, AnitoPlume was custom-built using a lean, native C++ open-source API stack consisting of OpenGL for graphics rendering, Eigen for numerical linear algebra, and dearIMGUI for the user interface layout.
This custom application architecture optimizes real-time performance on consumer-grade hardware.
By keeping the system dependencies minimal, the simulator successfully achieves real-time interactivity, running at an average rendering rate of 30 frames per second (FPS) on a mid-range system with 6 GB of video RAM (VRAM) \cite{del2026interactive}.

\subsection{Usability Design and Interactive Features}
The simulator also introduced several usability improvements over previous volcanic visualization systems, including simplified camera controls, interactive parameter editing, real-time plume direction tracking, and contextual information about nearby communities affected by volcanic activity.
These additions improved both accessibility and user experience while preserving interactive performance \cite{del2026interactive}.

\subsection{System Validation and Evaluation Metrics}
To evaluate the effectiveness of the simulator, \citeauthor{del2026interactive} conducted quantitative and qualitative assessments.
Visual fidelity was measured using the Structural Similarity Index (SSIM) and Root Mean Squared Error (RMSE) by comparing simulated eruption images with real-world photographs and surveillance footage of Taal Volcano.
The study reported high visual similarity, indicating that the synthesized plume closely resembled actual volcanic eruptions.
In addition, usability testing using the System Usability Scale (SUS) demonstrated that participants generally found the simulator easy to use and suitable for interactive exploration.

\subsection{Architectural Limitations and Future Trajectories}
Despite these accomplishments, the primary contribution of AnitoPlume lies in the design of its visualization framework rather than the underlying graphics API.
The study focused on plume simulation, terrain modeling, visual fidelity, and usability evaluation, while the OpenGL implementation served primarily as the rendering platform supporting these objectives.
Consequently, the original work did not investigate how the rendering architecture itself might influence performance, resource utilization, maintainability, or future extensibility.

The authors also acknowledged several opportunities for future development, including the incorporation of more advanced rendering techniques, expanded physics-based simulations, compute shader implementations, and additional GPU-accelerated features.
These recommendations highlight the need for a rendering architecture capable of supporting increasingly sophisticated graphics workloads while maintaining interactive performance.

\section{API Migration in Rendering Systems}
The migration of rendering software between graphics APIs is a well-known challenge in computer graphics. It often involves more than just replacing function calls. A successful port must account for differences in shader languages, resource-binding models, render-target handling, memory synchronization, and execution semantics. In practice, this means that a program cannot simply be translated line by line; instead, the rendering architecture must be reinterpreted in the context of the target API.

Previous work in graphics engineering has shown that API migration is most successful when the original system is structured in a modular way. Systems that separate rendering logic, scene management, and platform-specific code are often easier to port than tightly coupled implementations. This is particularly relevant for research software, where maintainability and clarity are essential. The process of porting from OpenGL to DirectX, therefore, becomes not only a technical exercise, but also an opportunity to improve the software architecture.

\subsection{Cross-API Shader Translation}
A major issue in porting visual effects from OpenGL to DirectX 12 is shader translation. OpenGL workflows usually rely on the OpenGL Shading Language (GLSL), while DirectX 12 expects the High-Level Shader Language (HLSL) and a more structured offline compilation pipeline. Existing developer practice tends to fall into two categories: manual shader rewriting for tight control over hardware-specific optimizations, or cross-compilation workflows that preserve shader intent while reducing rewrite cost \cite{Capodieci2022}. 

For a visually sensitive effect such as AnitoPlume's volcanic plumes, the second approach is highly attractive. Manual translation risks introducing numerical drift and subtle visual discrepancies in lighting and blending logic. This design philosophy is exemplified by modern research rendering frameworks like NVIDIA's Falcor, which architecturally separates program descriptions from backend compilation. By utilizing intermediate representations like SPIR-V, such engines prioritize shader portability over hard-coded backend assumptions. In a migration study, this supports the argument that accurate visual porting is not just about translating syntax from GLSL to HLSL, but about maintaining shader behavior across compilation targets using tool-assisted pipelines like the DirectX Shader Compiler (DXC) or SPIRV-Cross \cite{kinkor2022easy}.

\subsection{Pipeline State Management Adaptation}
A second major challenge is the shift from OpenGL’s dynamic state model to DirectX 12’s explicit Pipeline State Objects (PSOs). In architectures, OpenGL allows rendering states to be changed incrementally at runtime, with the graphics driver silently inferring and tracking state combinations \cite{Shiraef2016}. DirectX 12 completely eliminates this hidden driver-side management. Instead, explicit APIs require developers to bake the blend state, rasterizer state, depth-stencil state, primitive topology, and shader kernels into immutable PSOs ahead of time \cite{bossard2019high}. 

The necessity of this structural rewrite is evident in advanced rendering architectures like the Falcor framework. Rather than relying on hidden driver-side state inference, Falcor explicitly aggregates vertex layouts, shader kernels, rasterizer states, and framebuffer descriptions into a single, pre-compiled state object definition. At draw time, the render context binds this monolithic object directly before issuing the call. For plume rendering, adopting this explicit design is critical; even minor mismatches in depth testing or alpha blending across different PSOs can fundamentally alter how overlapping particle billboards accumulate and fade \cite{del2026interactive}. Therefore, PSO migration must be framed as a foundational architectural overhaul rather than a simple API substitution \cite{asplund2020directx}.

\subsection{Resource Synchronization and Memory Barriers}
The most technically demanding aspect of the explicit migration process is resource synchronization. In OpenGL, hardware hazards are concealed by the driver, meaning developers rarely need to explicitly manage the transitions between writing to and reading from the same texture or buffer \cite{Unterguggenberger2022}. DirectX 12, by contrast, transfers the burden of memory management entirely to the developer, requiring explicit resource barriers to ensure that GPU accesses occur in the correct deterministic order \cite{rossant2021high}. 

This explicit synchronization burden is concretely demonstrated in the Falcor engine, which utilizes a dedicated barrier system to route resources into their correct states before any updates, copies, or draw calls are executed. By manually handling transitions at both the resource and subresource levels, the architecture reinforces that state transitions are a central engine responsibility rather than an invisible driver service. For the AnitoPlume study, this supports a key argument: accurate migration is not just about producing a similar output image, but about guaranteeing that the GPU views each buffer in the correct state at the exact right microsecond. This is essential when updating thousands of particle billboards, as incorrect barrier placement during parallel draw calls leads to race conditions, stale texture reads, flickering artifacts, or catastrophic device crashes \cite{cheung2024applications}.
\begin{comment}
%
% IPR acknowledgement: the contents within this comment are from Ethel Ong's slides on RRL.
%
Guide on Writing your Related Works chapter
 
1. Identify the keywords with respect to your research
      One keyword = One document section
                Examples: 2.1 Story Generation Systems
			 2.2 Knowledge Representation

2.  Find references using these keywords

3.  For each of the references that you find,
        Check: Is it relevant to your research?
        Use their references to find more relevant works.

4. Identify a set of criteria for comparison.
       It will serve as a guide to help you focus on what to look for

5. Write a summary focusing on -
       What: A short description of the work
       How: A summary of the approach it utilized
       Findings: If applicable, provide the results
        Why: Relevance to your work

6. At the end of each section,  show a Table of Comparison of the related works 
   and your proposed project/system

\end{comment}
















