\chapter{Methodology}
\label{sec:methodology}
The primary objective of this research is to migrate and improve AnitoPlume's real-time simulation framework by modernizing its rendering pipeline, optimizing hardware utilization through new, explicit graphics APIs, and refining its underlying physical simulation models. 

\section{Systems Analysis}

We will investigate contemporary, explicit rendering paradigms to identify optimization opportunities within the existing legacy graphics pipeline (OpenGL). This involves studying low-level hardware abstraction techniques designed to reduce CPU overhead and improve memory bandwidth utilization. In parallel, existing computational fluid dynamics and particle-based simulation models will be evaluated to explore additional advanced algorithmic approaches to validate and potentially enhance the accuracy, scalability, and performance of the current physics implementations.

\section{Implementation and Migration}

The existing graphics architecture will be transitioned to a new, explicit graphics API. This pipeline re-engineering involves restructuring the rendering engine to utilize decoupled shading layers and unified abstractions. Concurrently, the underlying physical simulation logic, such as Lagrangian-based particle systems, will be validated and iteratively improved. Algorithmic refinements will be introduced to increase the fidelity of fluid-terrain interactions and volumetric behavior without compromising real-time performance constraints. Throughout this process, the system will undergo continuous integration and profiling. Optimization techniques will then be applied to resolve any bottlenecks in both graphics processing and physics computations.

\section{Usability Testing}

To verify that the architectural overhaul maintains or improves the interactive user experience, an exploratory usability test will be conducted with target beneficiaries:
\begin{itemize}
    \item Setup:  After screening and obtaining consent, participants will be given time to explore the interface.
    \item Task:  Participants will use the software to recreate specific Taal Volcano eruption plumes, following a set of predefined scenarios.
    \item Survey: Users will complete the System Usability Scale (SUS) questionnaire to rate the software's ease of use.
    \item Feedback: A closing interview will gather suggestions for improving the interface and overall performance.
\end{itemize}

\begin{table}[htbp]
\centering
\caption[Sample SUS questionnaire.]{Sample SUS questionnaire implemented by \cite{del2026interactive} 1 --- strongly disagree, 5 --- strongly agree.}\begin{tabular}{lp{12cm}}
\hline
Code & Question \\
\hline
Q1 & I think that I would use AnitoPlume frequently. \\
Q2 & I found AnitoPlume unnecessarily complex. \\
Q3 & I found AnitoPlume easy to use. \\
Q4 & I think that I would need the support of a technical person to be able to effectively and efficiently use AnitoPlume. \\
Q5 & I found that various features of the simulator were well integrated in AnitoPlume. \\
Q6 & I thought there was too much inconsistency in AnitoPlume. \\
Q7 & I would imagine that most people without prior experience in using similar software would easily learn how to use AnitoPlume. \\
Q8 & I found AnitoPlume very cumbersome (slow or complicated and therefore inefficient) to use. \\
Q9 & I felt very confident while using AnitoPlume. \\
Q10 & I needed to figure out a lot of things (such as hotkeys, navigating the UI) before I could confidently use AnitoPlume. \\
\hline
\end{tabular}
\end{table}


\section{Calendar of Activities}

For the proposal, a Gantt chart showing the schedule of the activities should be added. You can create a table like below, or create a figure. For example:

Table \ref{tab:timetableactivities} shows a Gantt chart of the activities.  Each bullet represents approximately one week worth of activity.

\begin{table}[htbp]
    \centering
    \caption{Proposed calendar of activities}
    \label{tab:calendaractivities}
    \begin{tabular}{l|c|c|c|c|c|c|c|c}
        \toprule
        \textbf{Activity} & \textbf{Aug} & \textbf{Sep} & \textbf{Oct} & \textbf{Nov} & \textbf{Dec} & \textbf{Jan} & \textbf{Feb} & \textbf{Mar} \\
        \midrule
        Literature review and reference refinement & X & X &  &  &  &  &  & \\
        Baseline analysis of AnitoPlume renderer &  & X & X &  &  &  &  & \\
        Environment setup and toolchain preparation &  &  & X & X &  &  &  & \\
        Rendering architecture design and migration planning &  &  &  & X & X &  &  & \\
        DirectX 12 implementation and shader porting &  &  &  &  & X & X & X & \\
        Resource synchronization and debugging &  &  &  &  &  & X & X & \\
        Performance evaluation and visual comparison &  &  &  &  &  &  & X & X \\
        Usability testing and thesis writing &  &  &  &  &  & X & X & X \\
        \bottomrule
    \end{tabular}
\end{table}



\begin{comment}

\begin{table}[]
    \centering
    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
         \toprule
         & \multicolumn{6}{c|}{2021} & \multicolumn{3}{c}{2022} \\
         \textbf{Activity} & \textbf{Apr} & \textbf{May} & \textbf{Jun} & \textbf{Jul} & \textbf{Aug} & \textbf{Sep} & \textbf{Jan} & \textbf{Feb} & \textbf{Mar} \\
         \midrule
         Data Collection & **** & **** & & & & & & & \\
         \midrule
         Software Design & & **** & **** & **** & **** & **** & & & \\
         \midrule
         Validation & & & & & & **** & **** & **** & \\
         \bottomrule
    \end{tabular}
    \caption{Gantt Chart of Activities}
    \label{tab:timetableactivities}
\end{table}
\end{comment}